Probe or Wait: Handling tail losses using Multipath TCP

### 2017-04-27 main.tex
Update

\begin{abstract}
Packet losses are known to affect the performance of latency sensitive applications in the Internet such as media streaming and gaming. Transport protocols recover from packet loss in order to provide reliable end to end communication and improving the quality of user experience. The efficiency of loss recovery greatly influences the completion time of flows. In this paper we focus on state of the art loss recovery mechanisms for TCP and Multipath TCP. We use controlled tail loss scenarios to evaluate the performance of loss recovery mechanisms. Based on the observations, we propose an enhancement to the tail loss recovery in Multipath TCP to improve the loss recovery time. Our experiment results show consistent end to end latency performance improvement in considered scenarios. 
\end{abstract}

@@

Multipath TCP~(MPTCP) is an experimental standard proposed as an extension to TCP~\cite{rfc6824}. It allows devices with multiple interfaces to transmit data simultaneously on both interfaces, thereby improving the throughput of end-to-end connections. Using MPTCP, a connection can have multiple TCP subflows using different interfaces on different routes. Each TCP subflow imparts the functional behavior of a standard TCP flow.

TCP has two mechanisms for detecting and recovering from packet losses: Fast Retransmit (FR) and Retransmission Timeout~(RTO)~\cite{Flach:2013}. Fast retransmit is triggered on receipt of a predecided number of duplicate acknowledgements, considering it as a loss indication. In case of RTO, a sender waits for the packet loss to trigger a timeout before retransmitting that packet. During timeout, the sender cannot send any packets, but with FR the out-of-order packets trigger duplicate acknowledgements. The sender can start retransmission of missing packets without waiting for timeout. For flows where there are sufficiently many packets to be transferred, FR is quicker than RTO in detecting and recovering packet losses. Fast retransmit coupled with fast recovery can send data outstanding when there is enough data to send in long flows. On the other hand, shorter flows more frequently depend on RTO for packet loss detection and recovery, thus incurring more latency. A single packet loss in a short flow may take many RTTs to detect and recover. This scenario is also applicable to the packets at the end of the flow (or tail), in a long flow. Although, RTO is reliable, even in tail loss~(losses at the end of transmission) recovery, it is not efficient in terms of recovery completion time. Tail losses are one of the major causes of end-to-end latency detoriation in modern networks~\cite{Flach:2013}.

TCP Loss Probe (TLP)~\cite{Flach:2013} is a mechanism to detect and recover from tail losses by employing a loss probe packet for a timeout shorter than the RTO. Probe timeout enables short flows to recover packet loss at the tail, by triggering FR. It assumes that other algorithms such as Early Retransmit~(ER)\cite{rfc5827} and Forward Acknowledgement~(FACK)~\cite{FACK} threshold based recovery are enabled. TLP is available in Linux kernel and enabled by default. Further, evaluations in~\cite{Rajiullah:2015} showed that TLP provides significant reductions in latency for short flows with one to n-degree tail losses.

 MPTCP connections start with one TCP flow to send data over, and may add more flows to the connection if configured to. MPTCP uses a connection level congestion window as well as subflow level congestion windows for each subflow. Data is split across the flows and each data packet is scheduled on one of the subflows based on minimum RTT. Packet losses in each subflow are detected and recovered in a similar fashion as that of TCP. Loss recovery can be conducted at different levels: the MPTCP connection level (meta level) and at subflow level. If the recovery is handled at meta level, a lost packet may be rescheduled and retransmitted on the available subflow with the lowest RTT. If recovery is handled at the subflow level, the packet must be retransmitted on the same subflow. An important part of loss recovery process at subflow level is to follow the TCP semantics and retransmit lost packets on the initial path of transfer. However, the MPTCP standard~\cite{rfc6824} does not restrict on, how the loss recovery should happen in the implementation and which subflow should retransmit the lost packets.

In this paper, we analyze the retransmission behavior of MPTCP in selected tail loss cases to identify the scope for improvement. We propose a change in the retransmission strategy of MPTCP to improve the latency performance. We provide related research on MPTCP retransmission in section~\ref{relwork}, and describe the experimental setup in section~\ref{exsetup}. We discuss observations from the experiments on state of the art retransmission behavior of TCP and MPTCP in section~\ref{disc}. Based on the observations, we propose an improvement to the current Linux implementation of TLP for MPTCP in section~\ref{impr}. In section~\ref{eval}, we provide an evaluation of the proposed TLP for MPTCP mechanism and conclude the paper in section~\ref{conc}. Our experimental results indicate improvement in the performance of TLP by reducing recovery time in the event of tail losses.

@@

Retransmission schemes are of significant importance in achieving low flow completion time. Several IETF RFCs such as SACK based recovery~\cite{rfc6675}, Limited Transmit~\cite{rfc3042} and Early Retransmission~\cite{rfc5827} have led to improved retransmission strategies for TCP. TLP is part of the various deployable loss recovery mechanisms proposed in~\cite{Flach:2013}, suggesting that current Linux TCP implementation provides room for improvement in several packet loss scenarios.

In a TCP flow, sequence numbers are used to identify lost packets and retransmit them. MPTCP uses two levels of sequence numbers to support efficient data transfer, namely data sequence and TCP sequence. Data sequence numbers are for the end-to-end data transfer and TCP sequence numbers are for individual flow data sequence. There is an association between data sequence numbers and TCP sequence numbers within a subflow.  A loss occurring in an individual TCP flow corresponds to a loss in end-to-end data. Multipath TCP, as an extension of TCP, uses the basics of TCP retransmission. However, the interaction with two levels makes the problem of retransmission more challenging than in TCP. The dual sequence numbering enables MPTCP to respond to loss of packets by retransmitting the lost packets on an alternate path.

@@

This paper analyzes the performance of state of the art loss recovery mechanism of MPTCP and presents a case for improvement in the tail loss scenario.For understanding the loss recovery and retransmission policies in MPTCP Linux implementation version 0.91, we consider several specific cases of tail losses. We use the standard settings of the Linux kernel. The Linux implementation of TCP is considered as the baseline TCP retransmission for implementation flexibility and availability of TLP support.

@@

Experiments use the CORE emulation platform~\cite{CORE} with a simple topology as depicted in Figure~\ref{fig1}. The client node connect to two wireless interfaces 3G or 4G and WLAN and the server node connect to a wired router.
CORE uses the Linux networking stack in each node of the topology by using a lightweight virtualization technique known as Linux network namespaces.
Characteristics of the connection setup and assumptions about the parameters are provided in Table~\ref{tab1}.
%The parameters are chosen to be close to real values of bandwidth and delay of respective technologies.
There is no attempt in this study to focus on the effect of link detailed characteristics in retransmission performance.

@@

%\kern-1em
\subsection{Testing retransmission with deterministic loss patterns}
In order to understand the retransmission behavior of the Linux MPTCP implementation and to reproduce the observed retransmissions, we use a deterministic drop pattern. Losses are generated by associating netem with corresponding interfaces and dropping select packets. This process is simplified by using the KAUNetem tool~\cite{Garcia2016}, which allows experimenter to e.g., drop packets based on their positions in a flow. A MPTCP connection starts with a single TCP subflow and subsequently, one or more subflows are added following an agreement between the client and the server. So one has to wait both in time and packets for the second subflow establishment. We send two bursts of 80 packets each and drop tail packets in the second burst at one of the interfaces. This is to ensure that the necessary and sufficient conditions for probe triggering are met and tail loss probe is generated on that subflow. We calculate the burst completion time for the second burst that has specified tail losses. The objective of this experiment is to evaluate retransmission mechanisms with tail loss (single packet loss and multiple tail loss) and temporary path loss. Three test cases are considered for experiments: single packet tail loss and two packet tail loss for tail loss evaluation and single packet along with probe~(in TLP) or retransmission~(in RTO) loss for path loss evaluation.

@@

\section{Observations and Discussion}\label{disc}
This section provides the performance analysis of TCP and MPTCP for the considered cases of single packet tail loss, two packet tail loss and retransmission or probe loss.
\subsection{TCP}
Retransmission behavior of TCP provides a base case for MPTCP performance as each individual MPTCP subflow is a TCP flow in itself. Our analysis
starts with the results using TCP and compare it with that of MPTCP. Figure~\ref{tcpf} represent the comparison of TCP 
burst completion times. Path asymmetry is irrelevant in this case as TCP uses a single path.

In the case of single packet tail loss as shown in Figure~\ref{t1p}, there is no difference in burst completion times of RTO and TLP settings. The retransmission
timeout and the probe timeout are the same as there is only one outstanding packet. This is a special case in the TLP
specification that adds 200\,ms to the PTO accommodating the case of delayed ACKs. The TLP implementation in Linux accounts
for the delayed ACK and waits 200\,ms before sending a probe. The minimum of PTO and RTO is considered for PTO. In the case of two packet tail loss as shown in Figure~\ref{t2p},
TLP performs better than the RTO retransmission for lower delay values. As the delay on the loss path increases RTO retransmission performs slightly better than TLP. In 120\,ms case, PTO is very close to RTO expiration and the second packet is recovered using ER which incurs additional delay in case of TLP. In RTO retransmission, all unacknowledged packets are retransmitted on a RTO expiration.

@@

The expected retransmission behavior of MPTCP for RTO retransmission and Linux default settings is shown 
in Figure.~\ref{mptiming}. The actual pattern might differ in cases, where the RTT is larger than 200\,ms. 
\renewcommand{\thesubfigure}{\roman{subfigure}}
\begin{figure*}[!tbp]
 \begin{subfigure}[b]{0.32\textwidth}
	\includegraphics[angle=0, width=\textwidth, natwidth=610, natheight=400]{images/timing1P.pdf}
	\caption{Single packet}\label{timing1P}
 \end{subfigure}
 \hfill
 \begin{subfigure}[b]{0.32\textwidth} 
	\includegraphics[angle=0, width=\textwidth, natwidth=610, natheight=400]{images/timing2P.pdf}
	\caption{Two packets }\label{timing2P}
 \end{subfigure} 
 \hfill
 \begin{subfigure}[b]{0.32\textwidth}
  	\includegraphics[angle=0, width=\textwidth, natwidth=610, natheight=400]{images/timing1PP.pdf}
	\caption{Single packet and probe loss}\label{timing1PP}
 \end{subfigure}
 \caption{Timing diagrams of MPTCP behavior on tail losses with a) RTO b) TLP }\label{mptiming}	
\end{figure*}

MPTCP uses an RTT based scheduler by default, that uses the lowest RTT path among the available paths to transmit
the data until its congestion window is full. Subsequently, the next lowest RTT path is chosen. MPTCP performance
is sensitive to asymmetry in paths. The burst completion times on single packet tail loss calculated for each delay 
configuration and retransmission setting is depicted in Figure~\ref{1p}. There is no difference in performance of RTO retransmission and TLP for 
cases when both paths are symmetric or the first path has lowest RTT. RTO performs better than TLP when the first path is not the lowest RTT path. In such
scenarios~(30\,ms-20\,ms,120\,ms-20\,ms), the RTO retransmission happens in the second path with lowest RTT following the behavior depicted
in Figure~\ref{timing1P}.

@@

In the case of single packet tail loss, the last outstanding packet is sent as probe and in case of n-degree tail loss,
the latest packet is sent as probe to trigger fast retransmit of the packets in between. In order to observe
the performance of n-degree tail loss, we consider dropping the last two packets. 
The results of two packet tail loss are shown in Figure~\ref{2p}. In this case, TLP performs better than the other 
two settings in all scenarios except when the packets are lost on a high delay interface. In this scenario, the MPTCP retransmission follows the behavior depicted in Figure~\ref{timing2P} and the delay difference is large enough that the retransmission of the last packet on the alternate path is faster than the probe on the primary path.

To further study the performance of TLP in the case when there is longer path interruption on one path, we tried to drop the probe packet along with the last packet. In this scenario, the RTO
setting results in much lower burst completion times. Comparing with the performance of TLP in TCP from Figure~\ref{t1pp}, where we saw that use of TLP improved the burst completion time, TLP in MPTCP did not improve the performance. The reason lies in the way the retransmission and loss probe transmission is carried out in MPTCP. Retransmission occurs after an RTO time on both paths.
Even if the retransmission fails on one path, the client receives the packet on the other path as shown in Figure~\ref{timing1PP}. For TLP, in
current implementation, the loss probe is sent on the actual path of the first transmission but not on both paths. This incomplete impartation of
TLP from TCP to MPTCP led to the increase in burst completion time by an RTO for MPTCP.

%\begin{table}[!ht]
%\centering
%\caption{Tail loss scenarios with tcp.early.retrans = 3 default}
%\label{ret3}
%\begin{tabular}{|l|l|l|l|l|}
%\hline
% testcase   & Asymmetry   & Metalevel          & Subflowlevel       &  \\\hline
%Tail drop 1 & 20ms - 30ms & rexmt in same path & rexmt in same path &  \\\hline
%Tail drop 2 & 30ms - 20ms & rexmt in same path & TBC                &  \\\hline
%Tail drop 3 & 20ms - 20ms & rexmt in same path & TBC                &  \\ \hline
%\end{tabular}
%\end{table}

%\begin{table}[!ht]
%\centering
%\caption{Tail loss scenarios with tcp.early.retrans = 0 ER disabled}
%\label{ret0}
%\begin{tabular}{|l|l|l|l|l|}
%\hline
% testcase   & Asymmetry   & Metalevel          & Subflowlevel       &  \\\hline
%Tail drop 1 & 20ms - 30ms & rexmt in same path & rexmt in same path &  \\\hline
%Tail drop 2 & 30ms - 20ms & rexmt in same path & TBC                &  \\\hline
%Tail drop 3 & 20ms - 20ms & rexmt in same path & TBC                & \\ \hline
%Tail drop 4 & 20ms - 120ms &  			&		&  \\ \hline
%Tail drop 5 & 120ms - 20ms &  			& 		& \\ \hline 
%\end{tabular}
%\end{table}

%\begin{table}[!ht]
%\centering
%\caption{Tail loss scenarios with tcp.early.retrans = 4 TLP disabled}
%\label{ret4}
%\begin{tabular}{|l|l|l|l|l|}
%\hline
% testcase   & Asymmetry   & Metalevel          & Subflowlevel       &  \\\hline
%Tail drop 1 & 20ms - 30ms & rexmt in same path & rexmt in same path &  \\\hline
%Tail drop 2 & 30ms - 20ms & rexmt in same path & TBC                &  \\\hline
%Tail drop 3 & 20ms - 20ms & rexmt in same path & TBC                & \\ \hline
%\end{tabular}
%\end{table}

%\kern-1em
\section{Improvements to TLP for MPTCP}\label{impr}
The goal of Tail Loss Probe~(TLP) is to reduce tail latency of short flows. It achieves this by converting retransmission timeouts (RTOs) occurring due to tail losses into fast recovery. TLP transmits one packet in two round-trips when a connection is in Open state and isn't receiving any ACKs. The transmitted packet, aka loss probe, can be either new or a retransmission. When there is tail loss, the ACK from a loss probe triggers FACK/early-retransmit based fast recovery, thus avoiding a costly retransmission timeout. Our results show that the current TLP implementation for MPTCP helps improving the latency performance in cases when the paths are symmetric or the first path~(for the packet) has lowest RTT. However, it reduces the performance in cases when the first path is not the lowest RTT path or when there is longer path interruption.

>>>>

\begin{abstract}
Packet losses are known to affect the performance of latency sensitive applications in the Internet such as media streaming and gaming. 
Transport protocols recover from packet loss in order to provide reliable end-to-end communication and improve user experience. The efficiency of 
loss recovery mechanisms influences the completion time of flows. In this paper we focus on state-of-the-art loss recovery mechanisms for TCP 
and Multipath TCP. We use controlled tail loss scenarios to evaluate the performance of loss recovery mechanisms, and based on the observations, 
we propose an enhanced tail loss recovery mechanism for Multipath TCP, to improve the loss recovery time. Our experiment results using Linux 
Multipath TCP implementation show consistent end to end latency performance improvement in considered scenarios. 
\end{abstract}

@@

Multipath TCP~(MPTCP)~\cite{rfc6824} is an experimental standard proposed as an extension to TCP. MPTCP allows devices with multiple interfaces to 
transmit data simultaneously on both interfaces, thereby improving the throughput of end-to-end connections. Using MPTCP, a connection can have 
multiple TCP subflows using different interfaces on different routes. Each TCP subflow imparts the functional behavior of a standard TCP flow.

TCP has two mechanisms for detecting and recovering from packet losses: Fast Retransmit (FR) and Retransmission Timeout~(RTO)~\cite{Flach:2013}. 
Fast retransmit is triggered on receipt of a predecided number of duplicate acknowledgements, considering it as an indication of loss. In the case 
of RTO, a sender waits for the packet loss to trigger a timeout before retransmitting packet. During such a timeout, the sender cannot send any 
packets. However, with FR the out-of-order packets trigger duplicate acknowledgements. The sender can start retransmission of missing packets 
without waiting for timeout. For flows where there are sufficiently many packets to be transferred, FR is quicker than RTO in detecting and 
recovering packet loss. Fast retransmit coupled with fast recovery can send data outstanding when there is enough data to send in long flows. 
On the other hand, shorter flows more frequently depend on RTO for packet loss detection and recovery, thus incurring more latency. A single 
packet loss in a short flow may take many RTTs to detect and recover. This scenario is also applicable to the packets at the end of the flow 
(or tail), of a long flow. Although RTO is reliable even in tail loss recovery, it is not efficient in terms of recovery completion time. Tail 
losses are one of the major causes of end-to-end latency detoriation in modern networks~\cite{Flach:2013}.

TCP Loss Probe (TLP)~\cite{Flach:2013} is a mechanism to detect and recover from tail losses by employing a loss probe packet for a timeout shorter
than the RTO. Probe timeout enables short flows to recover packet loss at the tail, by triggering FR. It assumes that other algorithms such as 
Early Retransmit~(ER)\cite{rfc5827} and Forward Acknowledgement~(FACK)~\cite{FACK} threshold based recovery are enabled. TLP is available in the Linux 
kernel and enabled by default. Further, evaluations in~\cite{Rajiullah:2015} show that TLP provides significant latency reductions for short 
flows with one to n-degree tail losses.

 MPTCP connections start with one TCP flow to send data over, and may add more flows to the connection if configured to. MPTCP uses a connection 
 level congestion window as well as subflow level congestion windows for each subflow. Data is split across the flows and each data packet is 
 scheduled on one of the subflows based on the minimum RTT. Packet losses in each subflow are detected and recovered in a similar fashion as that of 
 TCP. Loss recovery can be conducted at different levels: the MPTCP connection level (meta level) and at subflow level. If the recovery is handled 
 at meta level, a lost packet may be rescheduled and retransmitted on the available subflow with the lowest RTT. If recovery is handled at the 
 subflow level, the packet must be retransmitted on the same subflow. An important part of loss recovery process at the subflow level is to follow 
 the TCP semantics and retransmit lost packets on the initial path of transfer. However, the MPTCP standard~\cite{rfc6824} does not specify exactly, 
 how the loss recovery should happen in the implementation and which subflow should retransmit the lost packets.

In this paper, we analyze the retransmission behavior of MPTCP in selected tail loss cases to identify the scope for improvement. We propose a 
change in the retransmission strategy of MPTCP to improve the latency performance. We provide related research on MPTCP retransmission mechanisms in 
Section~\ref{relwork}, and describe the experimental setup in Section~\ref{exsetup}. We discuss observations from the experiments on state 
of the art retransmission behavior of TCP and MPTCP in Section~\ref{disc}. Based on the observations, we propose an improvement to the current 
Linux implementation of TLP for MPTCP in Section~\ref{impr}. In Section~\ref{eval}, we provide an evaluation of the proposed TLP for MPTCP 
mechanism and conclude the paper in Section~\ref{conc}. Our experimental results indicate improvement in the performance of TLP by reducing 
recovery time in the event of tail losses.

@@

Retransmission schemes are of significant importance in achieving a low flow completion time. Several IETF RFCs such as SACK based 
recovery~\cite{rfc6675}, Limited Transmit~\cite{rfc3042} and Early Retransmission~\cite{rfc5827} have led to improved retransmission 
strategies for TCP. TLP is part of the various deployable loss recovery mechanisms proposed in~\cite{Flach:2013}, suggesting that current 
Linux TCP implementation provides room for improvement in several packet loss scenarios.

In a TCP flow, sequence numbers are used to identify lost packets and retransmit them. MPTCP uses two levels of sequence numbers to support 
efficient data transfer, namely data sequence and TCP sequence. Data sequence numbers are for the end-to-end data transfer and TCP sequence 
numbers are for individual flow data sequence. There is an association between data sequence numbers and TCP sequence numbers within a subflow.  
A loss occurring in an individual TCP flow corresponds to a loss in end-to-end data. Multipath TCP, as an extension of TCP, uses the 
TCP retransmission mechanism. However, the interaction between the two levels makes the problem of retransmitting more challenging than in TCP. The dual sequence 
numbering enables MPTCP to respond to loss of packets by retransmitting the lost packets on an alternate path.

@@

This paper analyzes the performance of state-of-the-art loss recovery mechanism for MPTCP and presents a case for improvement in the tail loss 
scenario. For understanding the loss recovery and retransmission policies in the MPTCP Linux implementation (version 0.91), we consider several specific 
cases of tail losses. We use the default settings of the Linux kernel. The Linux implementation of TCP is considered as the baseline TCP 
retransmission for implementation flexibility and availability of TLP support.

@@

Experiments are run in the CORE emulation platform~\cite{CORE} with a simple topology as depicted in Figure~\ref{fig1}. The client node connect to two wireless interfaces 3G or 4G and WLAN and the server node connect to a wired router.
CORE uses the Linux networking stack in each node of the topology by using a lightweight virtualization technique known as Linux network namespaces.
Characteristics of the connection setup and assumptions about the parameters are provided in Table~\ref{tab1}.
There is no attempt in this study to focus on the effect of link specific characteristics in retransmission performance.

@@

%\kern-1em
\subsection{Testing retransmission mechanism with deterministic loss patterns}
In order to understand the retransmission behavior of the Linux MPTCP implementation and to reproduce the observed retransmissions, we use a 
deterministic drop pattern. Losses are generated by associating netem with corresponding interfaces and dropping select packets. This process 
is simplified by using the KAUNetem tool~\cite{Garcia2016}, which allows an experimenter to e.g., drop packets based on their positions in a flow. 
A MPTCP connection starts with a single TCP subflow and subsequently, one or more subflows are added following an agreement between the client 
and the server. So one has to wait both in time and packets for the second subflow establishment. We send two bursts of 80 packets each and drop 
tail packets in the second burst at one of the interfaces. This is to ensure that the necessary and sufficient conditions for probe triggering 
are met and tail loss probe is generated on that subflow. We calculate the burst completion time for the second burst that has specified tail 
losses. The objective of this experiment is to evaluate retransmission mechanisms with tail loss (single packet loss and multiple tail loss) and 
temporary path loss. Three test cases are considered for the experiments: single packet tail loss and two packet tail loss for tail loss evaluation 
and single packet along with probe~(in TLP) or retransmission~(in RTO) loss for path loss evaluation.

@@

\section{Observations and Discussion}\label{disc}
This section provides the performance analysis of TCP and MPTCP for the considered cases of single packet tail loss, two packet tail loss and 
retransmission or probe loss.
\subsection{TCP}
Retransmission behavior of TCP provides a base case for MPTCP performance as each individual MPTCP subflow is a TCP flow in itself. Our analysis
starts with the results using TCP and compares it with that of MPTCP. Figure~\ref{tcpf} represents the comparison of TCP burst completion times. Path
asymmetry is irrelevant in this case as TCP uses a single path.

In the case of single packet tail loss, as shown in Figure~\ref{t1p}, there is no difference in burst completion between RTO and TLP. The 
retransmission timeout and the probe timeout are the same as there is only one outstanding packet. This is a special case in the TLP
specification that adds 200\,ms to the PTO accommodating the case of delayed ACKs. The TLP implementation in Linux accounts for the delayed 
ACK and waits 200\,ms before sending a probe. The minimum of PTO and RTO is considered for PTO. In the case of two packet tail loss, as shown 
in Figure~\ref{t2p}, TLP performs better than the RTO retransmission for lower delay values. As the delay on the loss path increases RTO retransmission performs 
slightly better than TLP. In 120\,ms case, PTO is very close to RTO expiration and the second packet is recovered using ER which incurs additional 
delay in case of TLP. In RTO retransmission, all unacknowledged packets are retransmitted on a RTO expiration.

@@

The expected retransmission behavior of MPTCP for RTO retransmission and Linux default settings is shown 
in Figure~\ref{mptiming}. The actual pattern might be different in cases where the RTT is larger than 200\,ms.

\renewcommand{\thesubfigure}{\roman{subfigure}}
\begin{figure*}[!tbp]
 \begin{subfigure}[b]{0.32\textwidth}
	\includegraphics[angle=0, width=\textwidth, natwidth=610, natheight=400]{images/timing1P.pdf}
	\caption{Single packet}\label{timing1P}
 \end{subfigure}
 \hfill
 \begin{subfigure}[b]{0.32\textwidth} 
	\includegraphics[angle=0, width=\textwidth, natwidth=610, natheight=400]{images/timing2P.pdf}
	\caption{Two packets }\label{timing2P}
 \end{subfigure} 
 \hfill
 \begin{subfigure}[b]{0.32\textwidth}
  	\includegraphics[angle=0, width=\textwidth, natwidth=610, natheight=400]{images/timing1PP.pdf}
	\caption{Single packet and probe loss}\label{timing1PP}
 \end{subfigure}
 \caption{Timing diagrams of MPTCP behavior on tail losses with a) RTO b) TLP }\label{mptiming}	
\end{figure*}

MPTCP uses an RTT based scheduler by default, that uses the lowest RTT path among the available paths to transmit the data until its congestion 
window is full. Subsequently, the next lowest RTT path is chosen. MPTCP performance is sensitive to asymmetry in paths. The burst completion times 
on single packet tail loss calculated for each delay configuration and retransmission setting is depicted in Figure~\ref{1p}. There is no 
difference in performance between RTO retransmission and TLP for cases when both paths are symmetric or the first path has lowest RTT. RTO performs 
better than TLP when the first path is not the lowest RTT path. In such scenarios~(30\,ms-20\,ms,120\,ms-20\,ms), the RTO retransmission happens 
in the second path with lowest RTT following the behavior depicted in Figure~\ref{timing1P}.

@@

In the case of single packet tail loss, the last outstanding packet is sent as probe and in case of n-degree tail loss, the latest packet is sent 
as probe to trigger fast retransmit of the packets in between. In order to observe the performance of n-degree tail loss, we consider dropping the 
last two packets. The results of two packet tail loss are shown in Figure~\ref{2p}. In this case, TLP performs better than the other 
two settings in all scenarios except when packets are lost on a high delay interface. In this scenario, the MPTCP retransmission follows the 
behavior depicted in Figure~\ref{timing2P} and the delay difference is large enough that the retransmission of the last packet on the alternate 
path is faster than the probe on the primary path.

To further study the performance of TLP in the case when there is longer path interruption on one path, we tried to drop the probe packet along 
with the last packet. In this scenario, the RTO setting results in much lower burst completion times. Comparing with the performance of TLP in 
TCP from Figure~\ref{t1pp}, where we saw that use of TLP improved the burst completion time, TLP in MPTCP did not improve the performance. The 
reason lies in the way the retransmission and loss probe transmission is carried out in MPTCP. Retransmission occurs after an RTO time on both 
paths. Even if the retransmission fails on one path, the client receives the packet on the other path as shown in Figure~\ref{timing1PP}. For TLP, 
in the current implementation, the loss probe is sent on the actual path of the first transmission but not on both paths. This incomplete impartation 
of TLP from TCP to MPTCP led to the increase in burst completion time by an RTO for MPTCP.

%\kern-1em
\section{Improvements to TLP for MPTCP}\label{impr}
The goal of Tail Loss Probe~(TLP) is to reduce tail latency of short flows. It achieves this by converting retransmission timeouts (RTOs) occurring
due to tail losses into fast recovery. TLP transmits one packet in two round-trips when a connection is in Open state and is not receiving any ACKs.
The transmitted packet, or loss probe, can be either new or a retransmission. When there is tail loss, the ACK from a loss probe triggers 
FACK/early-retransmit based fast recovery, thus avoiding a costly retransmission timeout. Our results show that the current TLP implementation for 
MPTCP helps improving the latency performance in cases when the paths are symmetric or the first path~(for the packet) has lowest RTT. However, 
it reduces the performance in cases when the first path is not the lowest RTT path or when there is longer path interruption.
